Forecasting the 2024 U.S. Presidential Election

### 2024-11-04 paper/paper.qmd
Modify Prose over the paper

The 2024 U.S. Presidential Election, which will come to an end on Nov 5, 2024, draws the worldwide attention. Especially after Joe Biden dropped out, this cycle’s election is full of more uncertainties. Predicting the electoral result is crucial and meaningful for understanding potential shifts in voter behavior, with significant implications for future leadership and policy directions in the United States. Forecasting the election’s outcome becomes essential and meaningful for understanding potential shifts in the voters’ behavior. To make the prediction feasible, polling data becomes a primary tool for measuring voters’ sentiment and their expectations.

@@

The paper is organized as follows: @sec-data describes the dataset and the processing steps for model preparation. @sec-modelsetup explains the modeling approach, including predictor selection and the regression model structure. @sec-results presents the primary findings, detailing state-by-state predictions and overall electoral projections. @sec-discussion discusses the key findings, regional voting patterns, polling uncertainty and weakness. The @sec-Appendix1 provides n evaluation of the SurveyUSA polls and @sec-Appendix2 discusses about the idealized methodology on survey. Besides, model diagnostics and summary are also attached in @sec-model.

@@

This study uses R packages [@citeR] to clean and analyz the dataset, including libraries from tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], usmap [@usmap], corrplot [@corrplot], arrow [@arrow].

@@

- **Data Source and Quality Assurance**: The dataset is sourced from @fivethirty, a trusted platform recognized for its high standards in polling data quality. Only polls that meet specific criteria are included to ensure broad representation of likely U.S. Each poll documents essential details, including the pollster's identity, survey dates, sample size, and methodology. While these standards effectively capture respondents' voting intentions, they are based on historical benchmarks that may not fully adjust for recent changes in polling methods. The dataset also provides a numeric grade, transparency score, and pollster rating, calculated by @fivethirty using their internal criteria.
- **Temporal Limitations and Data Recency**: Polling data captures discrete snapshots of voter sentiment rather than continuous updates. Consequently, the analysis reflects intentions at particular times, which may not account for rapid shifts in opinion closer to Election Day. To address this, recency adjustments are applied to account for the time elapsed since each poll was conducted, helping to ensure the data remains reflective of current voter sentiment.
- **Participation and Response Bias**: Despite the rigorous methodology, certain biases in survey participation are difficult to avoid. Self-selection bias (where individuals choose whether to participate) and social desirability bias (where respondents may adjust their responses to appear socially acceptable) can affect data accuracy. These biases may introduce discrepancies between expressed opinions and actual voting behavior, potentially influencing forecast reliability.
- **Regional Sampling Imbalances**: The frequency of polling varies significantly by state, with battleground states receiving more frequent attention than states considered “safe.” This uneven distribution can lead to imbalances in regional representation, which may skew state-level and national forecasts. Such discrepancies highlight the challenge of producing a uniformly accurate forecast across all regions.

@@

The outcome variable of interest for this research is the percentage representing the level of public support for Donald Trump. The distribution is shown in @fig-pct, where support percentage for Donald Trump ranges from approximately 20% to 70%. The red dashed line at the 50% mark acts as a benchmark, indicating the threshold needed for majority support for winning.

@@

The Transparency Score measures how transparent a pollster is, calculated based on the amount of information disclosed about its polls, weighted by recency. The highest possible score is 10, while the lowest is 0. The distribution of Transparency Scores for the filtered pollsters shows a peak around 9 as presented in @fig-transparency-score, indicating that nine is the most common transparency score. @fig-transparency-score suggests that among the selected pollsters, there is a predominance of high transparency scores.

@@

The days until election represents the remaining time leading up to the final election day. This variable is calculated by finding the difference between the end date of each poll survey and the final election date, which is November 5, 2024.

@@

- Intercept ($\beta_0$): This is the predicted Trump polling percentage when all predictors (numeric grade, sample size, state, transparency score, and end date) are at their baseline or zero value.
- Numeric Grade ($\beta_1$): This coefficient measures how much Trump’s polling percentage changes as the pollster’s numeric grade increases. A positive and significant coefficient would indicate that higher-rated pollsters report better polling numbers for Trump, while a negative coefficient would suggest the opposite.
- Sample Size ($\beta_2$): This measures the impact of the number of respondents on Trump’s polling percentage. A positive coefficient would indicate that larger sample sizes are associated with higher polling percentages for Trump.
- State ($\beta_3$): The coefficients for the state variable represent differences in Trump’s polling percentage in each state compared to the reference state (baseline category). For example, if the coefficient for Florida is negative, it means Trump polls lower in Florida compared to the reference state. The state-level effects account for regional differences in Trump’s support. Some states may show significantly higher or lower levels of support, even after adjusting for the time of the poll and pollster quality.
- Transparency Score ($\beta_4$): This coefficient shows how much Trump’s polling percentage is affected by the transparency of the poll. A positive coefficient would indicate that polls with higher transparency tend to report higher polling percentages for Trump, whereas a negative coefficient would imply the opposite.
- Days Until Election ($\beta_5$): The counting down days is a time-related variable. A positive and significant coefficient would suggest that Trump’s polling percentage has decreased as the election date approaches, while a negative coefficient would suggest an increase in his polling percent over time.

@@

For future research, expanding the model to include demographic data, such as age, education level, and income could reveal a deeper analysis of voter behavior on a multi-sided scale. Additionally, real-time polling updates could offer a dynamic perspective, allowing the model to adapt to changes in public opinion as the election date approaches. This variability suggests that further studies could benefit from integrating real-time data and employing Bayesian methodologies to account for evolving voter preferences. Such an approach could improve the accuracy of predictions, especially in regions with historically unpredictable outcomes. These adjustments could improve model's reliability, particularly in closely contested states where the margin of error can significantly impact the final prediction.

@@

To address research questions, collecting data from every case in the target population is often impractical. Defining an appropriate target population, sampling frame, and sample size is essential to ensure accurate representation and feasible data collection , according to @taherdoost2016sampling. This is especially crucial in polling, where precise results rely on selecting a sample that reflects the broader population. Without clear definitions, polls risk producing biased or inaccurate outcomes.

@@

A sample is a subset of individuals or cases selected from the target population or sample frame to make inferences about that population [@fowler2013survey]. In the context of polls conducted by SurveyUSA, the sample refers to the survey respondents. Sample sizes vary across different SurveyUSA polls. For the 2024 U.S. presidential election cycle, SurveyUSA conducted 49 polls with sample sizes ranging from 507 to 2,330 registered or likely voters. The average sample size for these polls is approximately 1,045 households.

@@

SurveyUSA employs a mixed-method approach to recruitment, designed to reach a diverse and representative sample by incorporating online panels, telephone calls, and a text-to-web method. Some respondents are recruited using Random Digit Dialing (RDD) with telephone samples purchased from Aristotle. RDD generates phone numbers randomly, allowing SurveyUSA to reach both listed and unlisted numbers, thereby reducing bias associated with pre-existing contact lists. This technique is valuable for recruiting a broad demographic across various age groups, regions, and socioeconomic backgrounds for polls, such as voter preference surveys.

@@

* **Advantages:** The mixed approach offers several benefits. Simple random sampling ensures representativeness by giving all individuals an equal selection chance, a key feature that ensures representativeness and minimizes selection bias. Randomization helping to offset confounding effects from both known and unknown factors [@noor2022simple]. Convenience sampling through online panels further enhances data collection by making it faster and more cost-effective [@sedgwick2013convenience]. Additionally, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Rotating questions and answer choices helps mitigate bias, improving data reliability.\
* **Disadvantages:** Simple random sampling can be problematic for heterogeneous or dispersed populations [@noor2022simple]. Using phone-based data collection is time-consuming, as it requires contacting many individuals to reach an adequate response rate. Non-response issues, including busy signals, refusals, or hang-ups, are common in phone surveys. This can create non-response bias, where the views of non-participants differ from those of participants, compromising representativeness. The interviewer effect is another concern as respondents may adjust their answers in the presence of a live interviewer [@noor2022simple]. In addition, convenience sampling also presents challenges, primarily by introducing motivation bias. Participants’ willingness to join may depend on their interest in the survey topic, or dissatisfaction they want to voice, potentially skewing the sample’s representativeness [@sedgwick2013convenience].

@@

Addressing non-response in survey research is essential, as it occurs when selected participants do not participate or complete the survey, leading to non-response bias [@groves2012nonresponse]. This bias can compromise accuracy and representativeness by disrupting the information balance needed for the research question, particularly when certain groups are underrepresented, skewing findings away from the actual population [@sedgwick2013convenience]. For example, individuals who with demanding schedules, lower income or limited access to technology might be less likely to respond to a poll. Meanwhile, people with strong opinions or from specific demographic groups could be overrepresented in the results. As a result of non-response, the findings may misinform decision-makers, leading to conclusions that don’t truly represent the diverse views of the entire population. Addressing non-response is thus essential for producing balanced, reliable survey results that can genuinely inform policies, programs, and public opinion.

@@

SurveyUSA’s methodology reflects a balanced approach, leveraging various sampling approach and method to reach a representative sample. While its blend of probability and non-probability methods has strengths, such as cost-effectiveness and broad reach, it faces challenges related to telephone interview logistics, potential interviewer bias, and the limitations of fixed questionnaire options. Nevertheless, the inclusion of data weighting and question rotation adds credibility to its results, making SurveyUSA a reliable pollster for localized opinion research.

@@

## Sampling Strategy
Our sampling strategy combines **stratified** random sampling with **quota sampling** to ensure broad representation across key demographics. Stratified random sampling involves dividing the target population into subgroups (strata) based on demographic characteristics, then randomly sampling from each stratum. This technique enhances the representativeness of the sample and improves prediction accuracy by includes all important subpopulation [@taherdoost2016sampling] and reducing demographic and geographic sampling bias [@fivethirty].

@@

This stratification aligns with best practices in survey sampling, allowing for improved accuracy and generalizability in election forecasting by ensuring that subgroups are proportionately represented [@pewresearch]. However, in stratified sampling, selecting relevant stratification variables can be challenging, and it is neither feasible nor cost-effective to stratify across numerous variables simultaneously. For quota sampling, there is a risk of selection bias, and it does not guarantee the randomness, which may impact the sample’s representativeness.

@@

### Weighting
As discussed by @kalton2003weighting, weighting adjustments are frequently applied in surveys to address non- response and non-coverage, helping weighted sample estimates align with known population values. Thus, in this idealized methodology, post-stratification weights will be used to adjust for any oversampling or undersampling of specific demographic groups. For example, younger voters or underrepresented minorities will be weighted to reflect their actual proportions in the voting population. This adjustment ensures that the final sample closely mirrors the demographic composition of the target population, enhancing the accuracy and representativeness of the results.

@@

"Thank you for completing this survey. Your response has been recorded.
If you have any questions about this survey or would like to be informed about the results, please contact at [shaw.wei@mail.utoronto.ca](mailto:shaw.wei@mail.utoronto.ca)."

>>>>

The 2024 U.S. Presidential Election, which will come to an end on Nov 5, 2024, draws worldwide attention. Especially after Joe Biden dropped out, this cycle’s election is full of more uncertainties. Predicting the electoral result is crucial and meaningful for understanding potential shifts in voter behavior, with significant implications for future leadership and policy directions in the United States. Forecasting the election’s outcome becomes essential and meaningful for understanding potential shifts in the voters’ behavior. To make the prediction feasible, polling data becomes a primary tool for measuring voters’ sentiment and their expectations.

@@

The paper is organized as follows: @sec-data describes the dataset and the processing steps for model preparation. @sec-modelsetup explains the modeling approach, including predictor selection and the regression model structure. @sec-results presents the primary findings, detailing state-by-state predictions and overall electoral projections. @sec-discussion discusses the key findings, regional voting patterns, polling uncertainty and weakness. The @sec-Appendix1 provides n evaluation of the SurveyUSA polls and @sec-Appendix2 discusses about the idealized methodology of the survey. Besides, model diagnostics and summary are also attached in @sec-model.

@@

This study uses R packages [@citeR] to clean and analyze the dataset, including libraries from tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], usmap [@usmap], corrplot [@corrplot], arrow [@arrow].

@@

- **Data Source and Quality Assurance**: The dataset is sourced from @fivethirty, a trusted platform recognized for its high standards in polling data quality. Only polls that meet specific criteria are included to ensure a broad representation of likely U.S. Each poll documents essential details, including the pollster's identity, survey dates, sample size, and methodology. While these standards effectively capture respondents' voting intentions, they are based on historical benchmarks that may not fully adjust for recent changes in polling methods. The dataset also provides a numeric grade, transparency score, and pollster rating, calculated by @fivethirty using their internal criteria.
- **Temporal Limitations and Data Recency**: Polling data captures discrete snapshots of voter sentiment rather than continuous updates. Consequently, the analysis reflects intentions at particular times, which may not account for rapid shifts in opinion closer to Election Day. To address this, recency adjustments are applied to account for the time elapsed since each poll was conducted, helping to ensure the data remains reflective of current voter sentiment.
- **Participation and Response Bias**: Despite the rigorous methodology, certain biases in survey participation are difficult to avoid. Self-selection bias (where individuals choose whether to participate) and social desirability bias (where respondents may adjust their responses to appear socially acceptable) can affect data accuracy. These biases may introduce discrepancies between expressed opinions and actual voting behavior, potentially influencing forecast reliability.
- **Regional Sampling Imbalances**: The frequency of polling varies significantly by state, with battleground states receiving more frequent attention than states considered “safe.” This uneven distribution can lead to imbalances in regional representation, which may skew state-level and national forecasts. Such discrepancies highlight the challenge of producing a uniformly accurate forecast across all regions.

@@

The outcome variable of interest for this research is the percentage representing the level of public support for Donald Trump. The distribution is shown in @fig-pct, where the support percentage for Donald Trump ranges from approximately 20% to 70%. The red dashed line at the 50% mark acts as a benchmark, indicating the threshold needed for majority support for winning.

@@

The Transparency Score measures how transparent a pollster is, calculated based on the amount of information disclosed about its polls, weighted by recency. The highest possible score is 10, while the lowest is 0. The distribution of Transparency Scores for the filtered pollsters shows a peak around 9 as presented in @fig-transparency-score, indicating that 9 is the most common transparency score. @fig-transparency-score suggests that among the selected pollsters, there is a predominance of high transparency scores.

@@

The days until tge election represent the remaining time leading up to the final election day. This variable is calculated by finding the difference between the end date of each poll survey and the final election date, which is November 5, 2024.

@@

- Intercept ($\beta_0$): This is the predicted Trump polling percentage when all predictors (numeric grade, sample size, state, transparency score, and end date) are at their baseline or zero value.
- Numeric Grade ($\beta_1$): This coefficient measures how much Trump’s polling percentage changes as the pollster’s numeric grade increases. A positive and significant coefficient would indicate that higher-rated pollsters report better polling numbers for Trump, while a negative coefficient would suggest the opposite.
- Sample Size ($\beta_2$): This measures the impact of the number of respondents on Trump’s polling percentage. A positive coefficient would indicate that larger sample sizes are associated with higher polling percentages for Trump.
- State ($\beta_3$): The coefficients for the state variable represent differences in Trump’s polling percentage in each state compared to the reference state (baseline category). For example, if the coefficient for Florida is negative, it means Trump polls lower in Florida compared to the reference state. The state-level effects account for regional differences in Trump’s support. Some states may show significantly higher or lower levels of support, even after adjusting for the time of the poll and pollster quality.
- Transparency Score ($\beta_4$): This coefficient shows how much Trump’s polling percentage is affected by the transparency of the poll. A positive coefficient would indicate that polls with higher transparency tend to report higher polling percentages for Trump, whereas a negative coefficient would imply the opposite.
- Days Until Election ($\beta_5$): The counting down days is a time-related variable. A positive and significant coefficient would suggest that Trump’s polling percentage has decreased as the election date approaches, while a negative coefficient would suggest an increase in his polling percentage over time.

@@

For future research, expanding the model to include demographic data, such as age, education level, and income could reveal a deeper analysis of voter behavior on a multi-sided scale. Additionally, real-time polling updates could offer a dynamic perspective, allowing the model to adapt to changes in public opinion as the election date approaches. This variability suggests that further studies could benefit from integrating real-time data and employing Bayesian methodologies to account for evolving voter preferences. Such an approach could improve the accuracy of predictions, especially in regions with historically unpredictable outcomes. These adjustments could improve the model's reliability, particularly in closely contested states where the margin of error can significantly impact the final prediction.

@@

To address research questions, collecting data from every case in the target population is often impractical. Defining an appropriate target population, sampling frame, and sample size is essential to ensure accurate representation and feasible data collection, according to @taherdoost2016sampling. This is especially crucial in polling, where precise results rely on selecting a sample that reflects the broader population. Without clear definitions, polls risk producing biased or inaccurate outcomes.

@@

A sample is a subset of individuals or cases selected from the target population or sample frame to make inferences about that population [@fowler2013survey]. In the context of polls conducted by SurveyUSA, the sample refers to the survey respondents. Sample sizes vary across different SurveyUSA polls. For the 2024 U.S. presidential election cycle, SurveyUSA conducted 49 polls with sample sizes ranging from 507 to 2,330 registered or likely voters. The average sample size for these polls is approximately 1,045 households. Selecting a representative sample size helps ensure the survey results can be generalized to the target population while maintaining statistical precision [@fowler2013survey].

@@

SurveyUSA employs a mixed-method approach to recruitment, designed to reach a diverse and representative sample by incorporating online panels, telephone calls, and a text-to-web method. Some respondents were recruited using Random Digit Dialing (RDD) with telephone samples purchased from Aristotle. RDD generates phone numbers randomly, allowing SurveyUSA to reach both listed and unlisted numbers, thereby reducing bias associated with pre-existing contact lists. This technique is valuable for recruiting a broad demographic across various age groups, regions, and socioeconomic backgrounds for polls, such as voter preference surveys.

@@

* **Advantages:** The mixed approach offers several benefits. Simple random sampling ensures representativeness by giving all individuals an equal selection chance, a key feature that ensures representativeness and minimizes selection bias. Randomization helps to offset confounding effects from both known and unknown factors [@noor2022simple]. Convenience sampling through online panels further enhances data collection by making it faster and more cost-effective [@sedgwick2013convenience]. Additionally, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Rotating questions and answer choices helps mitigate bias, improving data reliability.\
* **Disadvantages:** Simple random sampling can be problematic for heterogeneous or dispersed populations [@noor2022simple]. Using phone-based data collection is time-consuming, as it requires contacting many individuals to reach an adequate response rate. Non-response issues, including busy signals, refusals, or hang-ups, are common in phone surveys. This can create non-response bias, where the views of non-participants differ from those of participants, compromising representativeness. The interviewer effect is another concern as respondents may adjust their answers in the presence of a live interviewer [@noor2022simple]. In addition, convenience sampling also presents challenges, primarily by introducing motivation bias. Participants’ willingness to join may depend on their interest in the survey topic, or dissatisfaction they want to voice, potentially skewing the sample’s representativeness [@sedgwick2013convenience].

@@

Addressing non-response in survey research is essential, as it occurs when selected participants do not participate or complete the survey, leading to non-response bias [@groves2012nonresponse]. This bias can compromise accuracy and representativeness by disrupting the information balance needed for the research question, particularly when certain groups are underrepresented, skewing findings away from the actual population [@sedgwick2013convenience]. For example, individuals with demanding schedules, lower income or limited access to technology might be less likely to respond to a poll. Meanwhile, people with strong opinions or from specific demographic groups could be overrepresented in the results. As a result of non-response, the findings may misinform decision-makers, leading to conclusions that don’t truly represent the diverse views of the entire population. Addressing non-response is thus essential for producing balanced, reliable survey results that can genuinely inform policies, programs, and public opinion.

@@

SurveyUSA’s methodology reflects a balanced approach, leveraging various sampling approaches and methods to reach a representative sample. While its blend of probability and non-probability methods has strengths, such as cost-effectiveness and broad reach, it faces challenges related to telephone interview logistics, potential interviewer bias, and the limitations of fixed questionnaire options. Nevertheless, the inclusion of data weighting and question rotation adds credibility to its results, making SurveyUSA a reliable pollster for localized opinion research.

@@

## Sampling Strategy
Our sampling strategy combines **stratified** random sampling with **quota sampling** to ensure broad representation across key demographics. Stratified random sampling involves dividing the target population into subgroups (strata) based on demographic characteristics, and then randomly sampling from each stratum. This technique enhances the representativeness of the sample and improves prediction accuracy by including all important subpopulations [@taherdoost2016sampling] and reducing demographic and geographic sampling bias [@fivethirty].

@@

This stratification aligns with best practices in survey sampling, allowing for improved accuracy and generalizability in election forecasting by ensuring that subgroups are proportionately represented [@pewresearch]. However, in stratified sampling, selecting relevant stratification variables can be challenging, and it is neither feasible nor cost-effective to stratify across numerous variables simultaneously. For quota sampling, there is a risk of selection bias, and it does not guarantee randomness, which may impact the sample’s representativeness.

@@

### Weighting
As discussed by @kalton2003weighting, weighting adjustments are frequently applied in surveys to address non-response and non-coverage, helping weighted sample estimates align with known population values. Thus, in this idealized methodology, post-stratification weights will be used to adjust for any oversampling or undersampling of specific demographic groups. For example, younger voters or underrepresented minorities will be weighted to reflect their actual proportions in the voting population. This adjustment ensures that the final sample closely mirrors the demographic composition of the target population, enhancing the accuracy and representativeness of the results.

@@

"Thank you for completing this survey. Your response has been recorded.
If you have any questions about this survey or would like to be informed about the results, please contact [shaw.wei@mail.utoronto.ca](mailto:shaw.wei@mail.utoronto.ca)."
